---
title: "Beer Market Change"
output: html_notebook
---

> How has the inputs, products and value of the brewing market changed over time?

```{r setup}
library(tidyverse)

brewing_materials <-
  readr::read_csv(
    'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv'
  ) %>%
  select(-month_prior_year,-ytd_current,-ytd_prior_year) %>%
  filter(!stringr::str_detect(type, "Total")) %>%
  mutate(year_month = lubridate::ymd(paste(year, month, 1)))

beer_taxed <-
  readr::read_csv(
    'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv'
  ) %>%
  select(-month_prior_year,-ytd_current,-ytd_prior_year) %>%
  filter(!stringr::str_detect(type, "Total")) %>%
  mutate(year_month = lubridate::ymd(paste(year, month, 1)))

brewer_size <-
  readr::read_csv(
    'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv'
  ) %>%
  filter(!stringr::str_detect(brewer_size, "Total")) %>%
  mutate(
    brewer_size =
      fct_relevel(
        as_factor(brewer_size),
        "Zero Barrels",
        "Under 1 Barrel",
        "1 to 1,000 Barrels",
        "1,001 to 7,500 Barrels",
        "7,501 to 15,000 Barrels",
        "15,001 to 30,000 Barrels",
        "30,001 to 60,000 Barrels",
        "60,001 to 100,000 Barrels",
        "100,001 to 500,000 Barrels",
        "500,001 to 1,000,000 Barrels",
        "1,000,000 to 6,000,000 Barrels",
        "6,000,001 Barrels and Over"
      )
  ) %>% 
  mutate(year = lubridate::ymd(paste(year, 1, 1)))
```

## Materials

```{r plot materials over time}
brewing_materials %>%
  ggplot(aes(
    x = year_month,
    y = month_current,
    group = type,
    colour = type
  )) +
  geom_line() +
  facet_wrap(. ~ type, scales = "free") +
  labs(title = "Brewing materials by month")
```

### Observations

* There is a sharp drop in the `month_current` in 2016 through to 2018
* There are two significant (anomalous?) spikes in `Hops (dry)`
* There is seasonality
* There are a number of (anomalous?) spikes in all the other metrics at various times, some of which seem co-incidental at the same same interval
  + the 2015 'drop'

## Taxed

```{r plot tax over time}
beer_taxed %>%
  ggplot(aes(
    x = year_month,
    y = month_current,
    group = type,
    colour = type
  )) +
  geom_line() +
  facet_wrap(. ~ type, scales = "free") +
  labs(title = "Brewing materials by month")
```

## Observations 

* There is no 'drop' in taxes to coincide with the materials
* There is a longer data set in taxes to 2020 in most cases
* There is only on siginifcant spike, and some trend changes

## Size

```{r plot size over time}
brewer_size %>%
  gather(key = "measure",
         value = 'value',
         -year,
         -brewer_size) %>%
  ggplot(aes(x = year, y = value, colour = brewer_size)) +
  geom_line() +
  scale_y_log10() +
  facet_wrap(. ~ measure, scales = "free") +
  labs(title = "Brewing size by year")
```

### Observations

* Log scale is required to see info between brewers of differetn scales on the same plot
* There is no _clear_ drop in 2016 either
  + There _might_ be something happening where brewers who produce `Under 1 Barrel` also suddenly drop in number in 2017, and drop in `taxable_removals` in 2019, but this is far from strong